---
title: "Making the resident and mutant single parent care strategies"
author: "Milly"
date: "2023-09-11"
output: pdf_document
---

# W6

Making a script that can invade any of the four strategies into each of the other strategies
Just break the chunks (remove ```{r}) for strategies you dont want to include in that trial
Need to also remove whatever you're turning into a continuous parameter 

##Baseline parameter values
```{r}
r0=5
K0=500

mu_m0=0.4
mu_f0=0.4

m_m0=0.6
m_f0=0.6

w_m0=0.4
w_f0=0.4

e_m0=0.5
e_f0=(1-e_m0)

a_m0= round((e_m0*(1-mu_m0)*m_m0), digits = 6)
a_f0= round((e_f0*(1-mu_f0)*m_f0), digits = 6)

e_mm0=a_m0
e_mf0=a_f0

e_sm0=e_m0*(1-mu_m0)
e_sf0=e_f0*(1-mu_f0)


e_mm0

e_sm0

a_m0
a_f0

e_sm0
e_sf0
```
##Resident:

Fecundity-investment trade-off
- (1-mu_m0) and (1-mu_f0) are proxies for level of initial egg investment, which trades-off with future reproduction
```{r}
rres=r0*exp(-((1-mu_m0)*e_m0 + (1-mu_f0)*e_f0))
rres
```

Resident is no care
```{r}

zres=6
c_mres=0.0
c_fres=0.0
c_tres=c_mres+c_fres
```

Resident 
```{r}
Kres=K0

mu_mres=mu_m0*exp(-zres*c_tres)
mu_fres=mu_f0*exp(-zres*c_tres)

m_mres=m_m0
m_fres=m_f0

e_mres=e_m0
e_fres=(1-e_mres)

w_mres=1-((1-w_m0)*exp(-c_mres))
w_fres=1-((1-w_f0)*exp(-(e_fres*(1-mu_f0)+(1-mu_m0)*e_mres+c_fres)))


a_mres= round((e_mres*(1-mu_mres)*m_mres), digits = 6)
a_fres= round((e_fres*(1-mu_fres)*m_fres), digits = 6)

e_mmres=a_mres
e_mfres=a_fres

e_smres=e_mres*(1-mu_mres)
e_sfres=e_fres*(1-mu_fres)


e_mres
e_fres

a_mres
a_fres

e_smres
e_sfres

```


Baseline for care
```{r}
z0=6

c_m0=0.35
c_f0=0.35
c_t0=c_m0+c_f0

```


Resident is female only care


zres=z0

c_mres=0
c_fres=c_f0 #0.05
c_tres=c_mres+c_fres


Kres=K0

e_mres=e_m0 #0.5
e_fres=(1-e_mres)

mu_mres=mu_m0*exp(-zres*c_tres)
mu_fres=mu_f0*exp(-zres*c_tres)

m_mres=m_m0
m_fres=m_f0

w_mres=w_m0
w_fres=1-((1-w_f0)*exp(-(e_fres*(1-mu_f0)+(1-mu_m0)*e_mres+c_fres)))


a_mres= round((e_mres*(1-mu_mres)*m_mres), digits = 6)
a_fres= round((e_fres*(1-mu_fres)*m_fres), digits = 6)

e_mmres=a_mres
e_mfres=a_fres

e_smres=e_mres*(1-mu_mres)
e_sfres=e_fres*(1-mu_fres)



Resident is male only care
``{r}
zres=z0

c_mres=c_m0 #0.05
c_fres=0
c_tres=c_mres+c_fres


Kres=K0

e_mres=e_m0 #0.5
e_fres=(1-e_mres)

mu_mres=mu_m0*exp(-zres*c_tres)
mu_fres=mu_f0*exp(-zres*c_tres)

m_mres=m_m0
m_fres=m_f0

w_mres=1-((1-w_m0)*exp(-c_mres))
w_fres=1-((1-w_f0)*exp(-(e_fres*(1-mu_f0)+(1-mu_m0)*e_mres)))
#took +c_f out, could've left it in and just let it be zer0? this would let it have extent of male/female care? - this is in biparental


a_mres= round((e_mres*(1-mu_mres)*m_mres), digits = 6)
a_fres= round((e_fres*(1-mu_fres)*m_fres), digits = 6)

e_mmres=a_mres
e_mfres=a_fres

e_smres=e_mres*(1-mu_mres)
e_sfres=e_fres*(1-mu_fres)



Resident is biparental care
``{r}
zres=z0

c_mres=c_m0 #0.05 can vary these here, not in the baseline bit
c_fres=c_f0 #0.05
c_tres=c_mres+c_fres


Kres=K0

e_mres=e_m0 #0.5
e_fres=(1-e_mres)

mu_mres=mu_m0*exp(-zres*c_tres)
mu_fres=mu_f0*exp(-zres*c_tres)

m_mres=m_m0
m_fres=m_f0

w_mres=1-((1-w_m0)*exp(-c_mres))
w_fres=1-((1-w_f0)*exp(-(e_fres*(1-mu_f0)+(1-mu_m0)*e_mres+c_fres)))
#includes +c_fres


a_mres= round((e_mres*(1-mu_mres)*m_mres), digits = 6)
a_fres= round((e_fres*(1-mu_fres)*m_fres), digits = 6)

e_mmres=a_mres
e_mfres=a_fres

e_smres=e_mres*(1-mu_mres)
e_sfres=e_fres*(1-mu_fres)





Equation for resident strategy
```{r}

AstarRes=round((Kres*(1-((w_mres*a_mres+w_fres*a_fres)/(e_mmres+e_mfres))*((mu_mres*e_mres+mu_fres*e_fres+m_mres*e_smres+m_fres*e_sfres)/(rres*a_fres)))), digits = 6)
AstarRes


AstarRound=round((K0*(1-((w_m0*a_m0+w_f0*a_f0)/(e_mm0+e_mf0))*((mu_m0*e_m0+mu_f0*e_f0+m_m0*e_sm0+m_f0*e_sf0)/(rres*a_f0)))), digits = 6)

AstarRound
```

##Mutant
    
    z0=10

    c_m0=0.05
    c_f0=0.05
    c_t0=c_m+c_f

Master code for care in mutant
```{r}
z=z0

c_m=c_m0
c_f=c_f0
c_t=c_m+c_f

```



Mutant is no care
- should still have fertility-initial investment trade-off (represented in r), should also have female adult death rate trade-off?
```{r}

K=K0

mu_m=mu_m0
mu_f=mu_f0

m_m=m_m0
m_f=m_f0

e_m=e_m0
e_f=1-e_m

w_m=w_m0
w_f=1-((1-w_f0)*exp(-(e_f*(1-mu_f0)+(1-mu_m0)*e_m)))

```


Mutant is female only care
``{r}

z= z0 #10

c_m=0
c_f=c_f0 #0.05
c_t=c_m+c_f

mu_m=mu_m0*exp(-z*c_t)
mu_f=mu_f0*exp(-z*c_t)

m_m=m_m0
m_f=m_f0

w_m=w_m0#w_m=1-((1-w_m0)*exp(-c_m))
w_f=1-((1-w_f0)*exp(-(e_f*(1-mu_f0)+(1-mu_m0)*e_m+c_f)))

```


Mutant is male only care
- in male only care there is still and initial egg investment trade-off in female adult death rate
``{r}

z= z0 #10

c_m=c_m0 #0.05
c_f=0
c_t=c_m+c_f

mu_m=mu_m0*exp(-z*c_t)
mu_f=mu_f0*exp(-z*c_t)

m_m=m_m0
m_f=m_f0

w_m=1-((1-w_m0)*exp(-c_m))
w_f=1-((1-w_f0)*exp(-(e_f*(1-mu_f0)+(1-mu_m0)*e_m)))


```

Mutant is biparental care
``{r}
z= z0 #10

c_m=c_m0 #0.05
c_f=c_f0 #0.05
c_t=c_m+c_f

mu_m=mu_m0*exp(-z*c_t)
mu_f=mu_f0*exp(-z*c_t)

m_m=m_m0
m_f=m_f0

w_m=1-((1-w_m0)*exp(-c_m))
w_f=1-((1-w_f0)*exp(-(e_f*(1-mu_f0)+(1-mu_m0)*e_m+c_f)))

```


Rest of mutant strategy
```{r}
mu_m=seq(0,1, length = 10000)


r=r0*exp(-((1-mu_m0)*e_m0 + (1-mu_f0)*e_f0))
#had to change from mu_m bc makes r a continuous parameter too ! problem in previous codes I'm guessing
#use e_m0 or e_m?

a_m=round((e_m*(1-mu_m)*m_m), digits = 6)
a_f=round((e_f*(1-mu_f)*m_f), digits = 6)


e_mm = a_m
e_mf = a_f


e_sm=e_m*(1-mu_m)
e_sf=e_f*(1-mu_f) 

```


Growth rate of mutant. 
- AstarRes represents the competition with the resident strategy it's invading. 
- invasion from rare assumes no intra-strategy competition
- using AstarRes
```{r}
b = mu_m*e_m  +  mu_f*e_f  + m_m*e_sm  + m_f*e_sf  +  w_m*a_m  +  a_f*w_f

c = (mu_m*e_m * w_m * a_m)   +   (mu_f*e_f *w_m * a_m)   +   (m_m*e_sm *w_m * a_m)  +  (m_f*e_sf *w_m * a_m)  +  (mu_m*e_m *a_f*w_f)  +  (mu_f*e_f *a_f*w_f)  +  (m_m*e_sm *a_f*w_f)  +  (m_f*e_sf *a_f*w_f)  - ((1-(AstarRes/K))* a_f *e_mm*r) - ((1-(AstarRes/K))* a_f *e_mf*r)

```



##Graph for the growth of mutant against a life history parameter of the strategy
- growth values above zero indicate invasion is possible. 
- use nlambda as the dominant eigen value (because is -b +sqrt)
```{r}
library(ggplot2)

plambda=-(b+sqrt(b^2-4*c))/2
nlambda=-(b-sqrt(b^2-4*c))/2


y_limits <- range(nlambda, na.rm = TRUE) 

library(scales)

data <- data.frame(mu_m = mu_m, nlambda = nlambda)

ggplot(data, aes(x = mu_m, y = nlambda)) +
  geom_line(size = 0.5, color = "red") +
  scale_x_continuous(breaks = pretty_breaks())+
  scale_y_continuous(breaks = pretty_breaks())+
   geom_hline(yintercept = 0, linetype = "solid", color = "blue", size = 0.2)+
  geom_vline(xintercept = 0, linetype = "solid", color = "blue", size = 0.2)+
  labs(x = "Male egg death rate (mu_m)", y = "Growth rate") +
  theme_minimal() + theme(axis.text = element_text(size = 9, face = "bold"))
```
  


  
  
  