---
title: "RK4-with delays"
output:
  pdf_document: default
  html_document: default
---

```{r}

pastvalue=function(dyno,lag){
  
tinc=0.01
bfl=3000

maxlag=(bfl-1)*tinc
p=0

if(lag>maxlag){
if(p==bfl) midpnt=1 else midpnt=p+1
past=hisy[midpnt,dyno]}

else

{
dispnt=floor((lag/tinc)+0.5)
disp=(dispnt*tinc-lag)/tinc

A=0.5*disp*(disp-1)
B=(1-disp^2)
C=0.5*disp*(disp+1)

if (dispnt>p) midpnt=p+dispnt else midpnt=bfl+p-dispnt

if(midpnt==1) leftpnt=bfl else leftpnt=midpnt-1

if(midpnt==bfl) rightpnt=1 else rightpnt=midpnt+1

R=hisy[rightpnt,dyno]
L=hisy[leftpnt,dyno]
M=hisy[midpnt,dyno]

past=A*L+B*M+C*R
}

past}
```

e_m = rate male eggs are produced
e_sm = rate of male eggs surviving the egg stage = e_m*(1-µ_m)

a_f is rate at which females enter the adult stage = e_f*(1-µ_f)*m_f*q_jf
q_jf = female juvenile survival rate
maybe do this but take out the juvenile survival bit

w_m = male adult death rate

\[
\frac{dE}{dt}= rA(t)a_f(1- \frac{A}{K}) - E(t)(µ_me_m + µ_fe_f + m_me_{sm} + m_fe_{sf})
\]

\[
\frac{dA}{dt}= (a_m + a_f)E(t) - (w_ma_m +w_fa_f)A(t)
\]

\[
a_f= e_f(1-µ_f)m_f
\]

\[
e_sf=e_f(1-mu_f)
\]

for us a_f = e_mf because we are not including juvenile surivival in a_f

y1 = E
y2 = A


trade-offs from 2013: wont type out bc so long


In this example there is 3 so nvar=3; yx[] is the derivative and y[] is the associated state variable:

```{r}
nvar=2

r0=6
K=250


mu_m0=0.5
mu_f0=0.5
w_m0=0.4
w_f0=0.4

e_m=0.5
e_f=(1-e_m)

a=1 # later z
c_m=0.35
c_f=0.35
c_t=c_m+c_f

mu_m=mu_m0#*exp(-a*c_t)
mu_f=mu_f0#*exp(-a*c_t)
w_m=1-((1-w_m0)*exp(-c_m))
w_f=1-((1-w_f0)*exp(-((1-mu_f0)*e_f+(1-mu_m0)*e_m+c_f)))


m_m0=0.5 #0.7 next
m_f0=0.5 #0.7

#Maturation rate increases with level and benefits of parental care
m_m=1-(1-m_m0)*exp(-a*c_t)
m_f=1-(1-m_f0)*exp(-a*c_t)

#Maturation rate increases with level of female initial egg investment
#m_m=1-((1-m_m0)*exp(-((1-mu_m0)*e_m)))
#m_f=1-((1-m_f0)*exp(-((1-mu_f0)*e_f)))
#m_m=m_m0
#m_f=m_f0

m_m1
m_m
m_f1
m_f


#r=r0*exp(-(((1-mu_m0)*e_m+(1-mu_f0)*e_f)))
#Adding in a /2 to try make range of r bigger
r=r0*exp(-(((1-mu_m0)*e_m+(1-mu_f0)*e_f))/2)

a_m=e_m*(1-mu_m)*m_m
a_f=e_f*(1-mu_f)*m_f

e_sm=e_m*(1-mu_m)
e_sf=e_f*(1-mu_f)



y=array(dim=(nvar))


derivs<-function(x,y){
	yx[1]=r*y[2]*a_f*(1-(y[2]/K))-y[1]*(mu_m*e_m + mu_f*e_f + m_m*e_sm + m_f*e_sf)
  yx[2]=(a_m+a_f)*y[1]-(w_m*a_m + w_f*a_f)*y[2]
	yx
}

```

We next set some initial model conditions, houskeeping variables associated with the number of iterations, integration step and buffer length, together with arrays we will use:

```{r}
#set the initial conditions 
y[1]=1
y[2]=1


#set the number of iterations (obs), integration step (h), buffer length (buflen) and start time (v)

obs=10000
buflen=3000
v=0.0
h=0.01

#arrays for the RK4 integrator (see below)

yt=array(1:nvar)
dyt=array(1:nvar)
dym=array(1:nvar)
yout=array(1:nvar)
yx=array(1:nvar)
eqn=array(1:nvar)

#these next two lines set up the history variables - it is necessary to carefully define what this history should be and ensure the state variable make sense

hisy=array(0,dim=c(buflen,nvar))
tmp=array(0,dim=c(buflen,nvar))
```

The next set of code defines, as a function, the RK4 integrator:

```{r}

#4th order runge-kutta function

rk4=function(dydx,y,t,n){
for(jj in	1:n)
	yt[jj]=y[jj]+dydx[jj]*h/2
	dyt=derivs(t+h/2,yt)
for(jj in 1:n)	
	yt[jj]=y[jj]+h/2*dyt[jj]
	dym=derivs(t+h/2,yt)
for(jj in 1:n)
	yt[jj]=y[jj]+h*dym[jj]
	dym=dyt+dym
	dyt=derivs(t+h,yt)
for(jj in 1:n)
	yout[jj]=y[jj]+(h/6)*(dydx[jj]+dyt[jj]+2*dym[jj])
	yout
}

```

The final set of code sets out the main part defining output arrays,  integrating the model and plotting the dynamics:

```{r}
#define output arrays
t1=array(1:obs)
x1=array(dim=c(obs,nvar))

#iterate the model 

for(i in 1:obs){
	eqn=derivs(v,y)
	y=rk4(eqn,y,v,nvar)

        #the next 7 lines of code update the history buffer

        for(ii in 1:buflen){
            for(jj in 1:nvar){
                tmp[ii,jj]=hisy[ii,jj]}}

        for(ii in 2:buflen){
            for(jj in 1:nvar){
                hisy[ii,jj]=tmp[ii-1,jj]}}

            for(jj in 1:nvar) hisy[1,jj]=y[jj]

	v=v+h
	x1[i,]=y
	t1[i]=v}

#plot the some results

plot(seq(0,max(t1),length=10),seq(0,max(x1),length=10),bty="l",xlab="Time",ylab="Numbers",main="Non-linear dynamics simulation",type="n")
lines(t1,x1[,1],col="red")	
lines(t1,x1[,2],col="blue")	
#lines(t1,x1[,3],col="green")	

```